---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Society Of Actuaries mortable web-crawler

<!-- badges: start -->
<!-- badges: end -->

The `SOAmort` is a `R` package that makes web crawling of a website call SOA mort. It has functions to obtain the mortality table data hosted on the website [mort.soa.org](https://mort.soa.org/).

## Install package

```{r eval=FALSE}
devtools::install_github("issactoast/SOAmort")
```

## Load package

```{r}
library(SOAmort)
```

The package consists of three functions:

- makeSearchInfo
- getTableInfo
- getTable

The mort.soa page provide the search options for the various tables using search parameters such as `Usage`, `Nation`, `Table layout` etc. The first two function will used for accessing the search result of the website.

### Search table using makeSearchInfo and getTableInfo

We will make a body to make a search request parameter which can be fed to the `getTableInfo` function. For example, if we want to search the table for USA's Annuitant Mortality select table, the search parameter will be the following: 

```{r}
myrequest <- makeSearchInfo(
    tableUsage = "Annuitant Mortality",
    nation = "United States of America",
    tableType = "Select")
```

After make the request parameter, we can feed this to the function `getTableInfo` as follows:

```{r}
result <- getTableInfo(myrequest)
```

result variable is a list which contains the infomation of the search result from the website.

```{r}
# Table numbers which satisfies to the request
result$TableIdentity

# Table Name/ Description
result$TableName

# Table Usage
result$ContentType
```

We can see there are 7 tables in the result.

### Access to the tables using table number

The `getTable` function allows you to access the actual table using the table number. For example, you can obtain the fist table 856 in the result variable as follows: 

```{r}
result <- SOAmort::getTable(856)
```

The package is using `httr` package to crawling, it will keep trying to catch the table utill it reach the maximum try or success to grab the table information.

```{r}
length(result)
```

Result shows that there are two table in the 856 table. To see the information about the table you can use `attributes` function.

```{r}
names(attributes(result[[1]]))
```

To see the difference between the two table, let us check the table description.

```{r}
attributes(result[[1]])$`Table Description`
attributes(result[[2]])$`Table Description`
```

It says the second table is the Railway Disabled Annuitants Mortality Ultimate Table whose minimum age 38 and the maximum age 95.

```{r}
RRBultmort <- result[[2]]
dim(RRBultmort)
min(RRBultmort$age)
max(RRBultmort$age)
head(RRBultmort)
```

```{r fig.align="center", out.width="70%"}
plot(RRBultmort$age, RRBultmort$`Column 1`,
     main = "The Railway Disabled Annuitants Mortality rate",
     xlab = "age",
     ylab = "mortality rate")
```

## Give me your feedback 

If you are having any issue and have an idea about the package, please make an issue in the [github repo](https://github.com/issactoast/SOAmort/issues).
